---
title: "Shaming"
author: "Harshil Kurapati"
format: html
execute:
  echo: false
---

```{r}
#| message: false
library(tidyverse)
library(primer.data)
library(tidymodels)
library(broom)
library(marginaleffects)
```

```{r}
#| cache: true
x <- shaming |>
  mutate(civ_engage = primary_00 + primary_02 + primary_04 +
                      general_00 + general_02 + general_04) |>
  select(primary_06, treatment, sex, age, civ_engage) |>
  drop_na() |>
  mutate(voter_class = factor(case_when(
    civ_engage %in% c(5, 6) ~ "Always Vote",
    civ_engage %in% c(3, 4) ~ "Sometimes Vote",
    civ_engage %in% c(1, 2) ~ "Rarely Vote"),
    levels = c("Rarely Vote", "Sometimes Vote", "Always Vote"))) |>
  mutate(voted = as.factor(primary_06)) |>
  drop_na()
```

```{r}
#| cache: true
fit_shaming <- logistic_reg(engine = "glm") |>
  fit(voted ~ age + sex + treatment*voter_class, data = x)
```

```{r}
plot_predictions(extract_fit_engine(fit_shaming),
                 condition = c("treatment", "voter_class"),
                 draw = FALSE) |>
  ggplot(aes(x = treatment, y = estimate, color = voter_class)) +
  geom_pointrange(aes(ymin = conf.low, ymax = conf.high), size = 0.7) +
  scale_x_discrete(limits = c("No Postcard", "Civic Duty", "Hawthorne",
                              "Self", "Neighbors")) +
  scale_y_continuous(labels = scales::label_percent(), limits = c(0, 0.75)) +
  facet_wrap(~ voter_class) +
  labs(
    title = "Predicted Turnout by Postcard and Voter Class",
    subtitle = "Neighbors moves Sometimes/Always voters most",
    x = NULL,
    y = NULL,
    color = "Voter Class",
    caption = "Source: Gerber, Green, Larimer 2008 via primer.data::shaming"
  ) +
  theme_minimal() +
  theme(legend.position = "none", axis.text.x = element_text(angle = 30, hjust = 1))
```

Voter turnout in low-salience primary elections is highly sensitive to social pressure tactics. Using Gerber, Green, and Larimer's 2008 Michigan field experiment with about 344,000 registered voters we estimate the average causal effect of the Neighbors postcard relative to no postcard on the probability of voting in the August 2006 primary. One concern is that this experiment was conducted in Michigan in 2006 and political messaging norms have changed significantly since then so the persuasive power of social pressure postcards on a 2026 Texas electorate might be quite different. We model the probability of voting using logistic regression as a function of age, sex, treatment, and voter class with an interaction between treatment and voter class to allow the postcard effect to vary by past voting history. We estimate that the Neighbors postcard raises the probability of voting by about 8.1 percentage points on average with the effect concentrated among Sometimes and Always voters at 8 to 9 percentage points and notably smaller among Rarely Vote voters at 4 percentage points.
